---
title: "Exploratory Data Analysis"
format: html
editor: visual
---

```{r}
library(haven)
library(tidyverse)
library(gridExtra)
library(grid)
library(gridtext)
```

```{r}
sleep <- read_xpt("SLQ_J.XPT")
diet_intake <- read_xpt("DR1IFF_J.XPT")
diet_total <- read_xpt("DR1TOT_J.XPT")
```

```{r}
dat <- merge(sleep, diet_total, by = "SEQN")

dat2 <- merge(sleep, diet_intake, by = "SEQN")
dat2_breakfast <- dat2 |> filter(DR1_030Z == 1)
dat2_lunch <- dat2 |> filter(DR1_030Z == 2)
dat2_dinner <- dat2 |> filter(DR1_030Z == 3)
```

# Univariate Sleep Data
```{r}
sleep |>
  ggplot(aes(x = SLQ120)) +
  geom_bar() +
  xlab("Feeling Overly Sleepy During the Day")

sleep |>
  ggplot(aes(x = SLQ050)) +
  geom_bar() +
  xlab("Ever Told the Doctor You Had Trouble Sleeping")
```

# Macronutrient Consumption vs. Sleep Quality
## Macronutrient Consumption vs. Weekday Sleep
```{r}
prot_weekday <- dat |>
  ggplot(aes(x = DR1TPROT, y = SLD012)) +
  geom_jitter(alpha = 0.7) +
  ylab("")

summary(lm(SLD012 ~ DR1TPROT, data = dat)) # significant, -

carb_weekday <- dat |>
  ggplot(aes(x = DR1TCARB, y = SLD012)) +
  geom_jitter(alpha = 0.7) +
  ylab("")

summary(lm(SLD012 ~ DR1TCARB, data = dat)) # significant, -

fat_weekday <- dat |>
  ggplot(aes(x = DR1TTFAT, y = SLD012)) +
  geom_jitter(alpha = 0.7) +
  ylab("")

summary(lm(SLD012 ~ DR1TTFAT, data = dat)) # significant, -

fibe_weekday <- dat |>
  ggplot(aes(x = DR1TFIBE, y = SLD012)) +
  geom_jitter(alpha = 0.7) +
  ylab("")

summary(lm(SLD012 ~ DR1TFIBE, data = dat)) # not significant


yleft_weekday = richtext_grob("Hours of Sleep on Weekdays", rot=90)

top_weekday = richtext_grob("Macronutrient Consumption vs. Weekday Sleep")

grid.arrange(prot_weekday, carb_weekday, fat_weekday, fibe_weekday, ncol=2, nrow =2, left = yleft_weekday, top = top_weekday)
```

## Macronutrient Consumption vs. Weekend Sleep
```{r}
prot_weekend <- dat |>
  ggplot(aes(x = DR1TPROT, y = SLD013)) +
  geom_jitter(alpha = 0.7) +
  ylab("")

summary(lm(SLD013 ~ DR1TPROT, data = dat)) # not significant

carb_weekend <- dat |>
  ggplot(aes(x = DR1TCARB, y = SLD013)) +
  geom_jitter(alpha = 0.7) +
  ylab("")

summary(lm(SLD013 ~ DR1TCARB, data = dat)) # not significant

fat_weekend <- dat |>
  ggplot(aes(x = DR1TTFAT, y = SLD013)) +
  geom_jitter(alpha = 0.7) +
  ylab("")

summary(lm(SLD013 ~ DR1TTFAT, data = dat)) # not significant

fibe_weekend <- dat |>
  ggplot(aes(x = DR1TFIBE, y = SLD013)) +
  geom_jitter(alpha = 0.7) +
  ylab("")

summary(lm(SLD013 ~ DR1TFIBE, data = dat)) # not significant

yleft_weekend = richtext_grob("Hours of Sleep on Weekends", rot=90)

top_weekend = richtext_grob("Macronutrient Consumption vs. Weekend Sleep")

grid.arrange(prot_weekend, carb_weekend, fat_weekend, fibe_weekend, ncol=2, nrow =2, left = yleft_weekend, top = top_weekend)
```

## Macronutrient Consumption vs. Feeling Overly Sleepy During the Day
```{r}
prot_sleepy <- dat |>
  mutate(SLQ120 = as.factor(SLQ120)) |>
  ggplot(aes(x = DR1TPROT, y = SLQ120)) +
  geom_boxplot() +
  ylab("")

summary(lm(SLQ120 ~ DR1TPROT, data = dat)) # not significant

carb_sleepy <- dat |>
  mutate(SLQ120 = as.factor(SLQ120)) |>
  ggplot(aes(x = DR1TCARB, y = SLQ120)) +
  geom_boxplot() +
  ylab("")

summary(lm(SLQ120 ~ DR1TCARB, data = dat)) # significant, +

fat_sleepy <- dat |>
  mutate(SLQ120 = as.factor(SLQ120)) |>
  ggplot(aes(x = DR1TTFAT, y = SLQ120)) +
  geom_boxplot() +
  ylab("")

summary(lm(SLQ120 ~ DR1TTFAT, data = dat)) # significant, +

fibe_sleepy <- dat |>
  mutate(SLQ120 = as.factor(SLQ120)) |>
  ggplot(aes(x = DR1TFIBE, y = SLQ120)) +
  geom_boxplot() +
  ylab("")

summary(lm(SLQ120 ~ DR1TFIBE, data = dat)) # significant, -

yleft_sleepy = richtext_grob("Feeling Overly Sleepy During the Day", rot=90)

top_sleepy = richtext_grob("Macronutrient Consumption vs. Feeling Overly Sleepy")

grid.arrange(prot_sleepy, carb_sleepy, fat_sleepy, fibe_sleepy, ncol=2, nrow =2, left = yleft_sleepy, top = top_sleepy)
```

## Macronutrient Consumption vs. Trouble Sleeping
```{r}
prot_trouble <- dat |>
  mutate(SLQ050 = as.factor(SLQ050)) |>
  ggplot(aes(x = DR1TPROT, y = SLQ050)) +
  geom_boxplot() +
  ylab("")

summary(lm(SLQ050 ~ DR1TPROT, data = dat)) # not significant

carb_trouble <- dat |>
  mutate(SLQ050 = as.factor(SLQ050)) |>
  ggplot(aes(x = DR1TCARB, y = SLQ050)) +
  geom_boxplot() +
  ylab("")

summary(lm(SLQ050 ~ DR1TCARB, data = dat)) # significant, +

fat_trouble <- dat |>
  mutate(SLQ050 = as.factor(SLQ050)) |>
  ggplot(aes(x = DR1TTFAT, y = SLQ050)) +
  geom_boxplot() +
  ylab("")

summary(lm(SLQ050 ~ DR1TTFAT, data = dat)) # not significant

fibe_trouble <- dat |>
  mutate(SLQ050 = as.factor(SLQ050)) |>
  ggplot(aes(x = DR1TFIBE, y = SLQ050)) +
  geom_boxplot() +
  ylab("")

summary(lm(SLQ050 ~ DR1TFIBE, data = dat)) # significant, +

yleft_trouble = richtext_grob("Told Doctor You Had Trouble Sleeping", rot=90)

top_trouble = richtext_grob("Macronutrient Consumption vs. Trouble Sleeping")

grid.arrange(prot_trouble, carb_trouble, fat_trouble, fibe_trouble, ncol=2, nrow =2, left = yleft_trouble, top = top_trouble)
```

# Temporal Eating Habits vs. Sleep Quality
## Meal Time vs. Weekday Sleep
```{r}
breakfast_weekday <- dat2 |>
  filter(DR1_030Z == 1) |>
  ggplot(aes(x = DR1_020, y = SLD012)) +
  geom_point() +
  ylab("") +
  xlab("Breakfast Time")

summary(lm(SLD012 ~ DR1_020, data = dat2_breakfast)) # significant, +

lunch_weekday <- dat2 |>
  filter(DR1_030Z == 2) |>
  ggplot(aes(x = DR1_020, y = SLD012)) +
  geom_point() +
  ylab("") +
  xlab("Lunch Time")

summary(lm(SLD012 ~ DR1_020, data = dat2_lunch)) # significant, +

dinner_weekday <- dat2 |>
  filter(DR1_030Z == 3) |>
  ggplot(aes(x = DR1_020, y = SLD012)) +
  geom_point() +
  ylab("") +
  xlab("Dinner Time")

summary(lm(SLD012 ~ DR1_020, data = dat2_dinner)) # not significant

yleft_meal_weekday = richtext_grob("Hours of Sleep on Weekdays", rot=90)

top_meal_weekday = richtext_grob("Meal Time vs. Weekday Sleep")

grid.arrange(breakfast_weekday, lunch_weekday, dinner_weekday, ncol=2, nrow =2, left = yleft_meal_weekday, top = top_meal_weekday)
```

## Meal Time vs. Weekend Sleep
```{r}
breakfast_weekend <- dat2 |>
  filter(DR1_030Z == 1) |>
  ggplot(aes(x = DR1_020, y = SLD013)) +
  geom_point() +
  ylab("") +
  xlab("Breakfast Time")

summary(lm(SLD013 ~ DR1_020, data = dat2_breakfast)) # significant, +

lunch_weekend <- dat2 |>
  filter(DR1_030Z == 2) |>
  ggplot(aes(x = DR1_020, y = SLD013)) +
  geom_point() +
  ylab("") +
  xlab("Lunch Time")

summary(lm(SLD013 ~ DR1_020, data = dat2_lunch)) # significant, +

dinner_weekend <- dat2 |>
  filter(DR1_030Z == 3) |>
  ggplot(aes(x = DR1_020, y = SLD013)) +
  geom_point() +
  ylab("") +
  xlab("Dinner Time")

summary(lm(SLD013 ~ DR1_020, data = dat2_dinner)) # not significant

yleft_meal_weekend = richtext_grob("Hours of Sleep on Weekends", rot=90)

top_meal_weekend = richtext_grob("Meal Time vs. Weekend Sleep")

grid.arrange(breakfast_weekend, lunch_weekend, dinner_weekend, ncol=2, nrow =2, left = yleft_meal_weekend, top = top_meal_weekend)
```

## Meal Time vs. Feeling Overly Sleepy During the Day
```{r}
breakfast_sleepy <- dat2 |>
  filter(DR1_030Z == 1) |>
  mutate(SLQ120 = as.factor(SLQ120)) |>
  ggplot(aes(x = DR1_020, y = SLQ120)) +
  geom_boxplot() +
  ylab("") +
  xlab("Breakfast Time")

summary(lm(SLQ120 ~ DR1_020, data = dat2_breakfast)) # significant, +

lunch_sleepy <- dat2 |>
  filter(DR1_030Z == 2) |>
  mutate(SLQ120 = as.factor(SLQ120)) |>
  ggplot(aes(x = DR1_020, y = SLQ120)) +
  geom_boxplot() +
  ylab("") +
  xlab("Lunch Time")

summary(lm(SLQ120 ~ DR1_020, data = dat2_lunch)) # not significant

dinner_sleepy <- dat2 |>
  filter(DR1_030Z == 3) |>
  mutate(SLQ120 = as.factor(SLQ120)) |>
  ggplot(aes(x = DR1_020, y = SLQ120)) +
  geom_boxplot() +
  ylab("") +
  xlab("Dinner Time")

summary(lm(SLQ120 ~ DR1_020, data = dat2_dinner)) # significant, -

yleft_meal_sleepy = richtext_grob("Feeling Overly Sleepy During the Day", rot=90)

top_meal_sleepy = richtext_grob("Meal Time vs. Feeling Overly Sleepy")

grid.arrange(breakfast_sleepy, lunch_sleepy, dinner_sleepy, ncol=2, nrow =2, left = yleft_meal_sleepy, top = top_meal_sleepy)
```

## Meal Time vs. Trouble Sleeping
```{r}
breakfast_trouble <- dat2 |>
  filter(DR1_030Z == 1) |>
  mutate(SLQ050 = as.factor(SLQ050)) |>
  ggplot(aes(x = DR1_020, y = SLQ050)) +
  geom_boxplot() +
  ylab("") +
  xlab("Breakfast Time")

summary(lm(SLQ050 ~ DR1_020, data = dat2_breakfast)) # significant, +

lunch_trouble <- dat2 |>
  filter(DR1_030Z == 2) |>
  mutate(SLQ050 = as.factor(SLQ050)) |>
  ggplot(aes(x = DR1_020, y = SLQ050)) +
  geom_boxplot() +
  ylab("") +
  xlab("Lunch Time")

summary(lm(SLQ050 ~ DR1_020, data = dat2_lunch)) # not significant

dinner_trouble <- dat2 |>
  filter(DR1_030Z == 3) |>
  mutate(SLQ050 = as.factor(SLQ050)) |>
  ggplot(aes(x = DR1_020, y = SLQ050)) +
  geom_boxplot() +
  ylab("") +
  xlab("Dinner Time")

summary(lm(SLQ050 ~ DR1_020, data = dat2_dinner)) # significant, +

yleft_meal_trouble = richtext_grob("Told Doctor You Had Trouble Sleeping", rot=90)

top_meal_trouble = richtext_grob("Meal Time vs. Trouble Sleeping")

grid.arrange(breakfast_trouble, lunch_trouble, dinner_trouble, ncol=2, nrow =2, left = yleft_meal_trouble, top = top_meal_trouble)
```

# Number of Hours Awake vs. Sleep Quality
```{r}
awake <- sleep |>
  mutate(hours_awake_weekdays = 24 - SLD012,
         hours_awake_weekends = 24 - SLD013)

awake |>
  mutate(SLQ120 = as.factor(SLQ120)) |>
  ggplot(aes(x = hours_awake_weekdays, y = SLQ120)) +
  geom_boxplot() +
  ylab("Feeling Overly Sleepy During the Day") +
  xlab("Hours Awake on Weekdays")

summary(lm(SLQ120 ~ hours_awake_weekdays, data = awake)) # not significant

awake |>
  mutate(SLQ050 = as.factor(SLQ050)) |>
  ggplot(aes(x = hours_awake_weekdays, y = SLQ050)) +
  geom_boxplot() +
  ylab("Ever Told the Doctor You Had Trouble Sleeping") +
  xlab("Hours Awake on Weekdays")

summary(lm(SLQ050 ~ hours_awake_weekdays, data = awake)) # significant, -

awake |>
  mutate(SLQ120 = as.factor(SLQ120)) |>
  ggplot(aes(x = hours_awake_weekends, y = SLQ120)) +
  geom_boxplot() +
  ylab("Feeling Overly Sleepy During the Day") +
  xlab("Hours Awake on Weekends")

summary(lm(SLQ120 ~ hours_awake_weekends, data = awake)) # not significant

awake |>
  mutate(SLQ050 = as.factor(SLQ050)) |>
  ggplot(aes(x = hours_awake_weekends, y = SLQ050)) +
  geom_boxplot() +
  ylab("Ever Told the Doctor You Had Trouble Sleeping") +
  xlab("Hours Awake on Weekends")

summary(lm(SLQ050 ~ hours_awake_weekends, data = awake)) # significant, -
```
